Digital Analysis of Paintings

### 2014-05-03 paper.tex
Add paper 1.

@@

\newacronym{APICSS}{APICSS}{Art Painting Image Color Semantics}
\newacronym{RGB}{RGB}{Red, Green, Blue}
\newacronym{HSL}{HSL}{Hue, Saturation, Luminance}
\newacronym{HSV}{HSV}{Hue, Saturation, Value}
\newacronym{PCA}{PCA}{Principal Component Analysis}
\newacronym{t-SNE}{t-SNE}{t-Distributed Stochastic Neighbor Embedding}
\newacronym{MDS}{MDS}{Multidimensional Scaling}

@@

@@

>>>>

\usepackage{paralist}

@@

\newacronym{APICSS}{APICSS}{Art Painting Image Color Semantics}
\newacronym{RGB}{RGB}{Red, Green, Blue}
\newacronym{HSL}{HSL}{Hue, Saturation, Luminance}
\newacronym{HSV}{HSV}{Hue, Saturation, Value}
\newacronym{PCA}{PCA}{Principal Component Analysis}
\newacronym{t-SNE}{t-SNE}{t-Distributed Stochastic Neighbor Embedding}
\newacronym{MDS}{MDS}{Multidimensional Scaling}
\newacronym{SVM}{SVM}{Support Vector Machine}

@@

To be able to extract brushstrokes it is important to understand how an expert
might go about the task. A lot of analysis techniques tend to be inspired from
how a human might go about the task, this also help a human to understand how
the process was gone about and may give extra meaning to the results.

There is a lot of knowledge to be gained from looking at the expert process,
and this knowledge can be applied to reduce noise and parameters from the
extraction process. The knowledge of painting materials, the state of the
preservation of the painting and the working methods of the artist themselves.

Interestingly, this leads to the conclusion that parts of the painting should
actually be discounted from the analysis, elements which have deteriorated over
time due to instabilities in the painting materials or the lack of preservation
of the painting leads to what most computer vision researchers would call
``noise''. Certain areas of the painting may not have been painting by the
artist's hand, such as cracks in the painting from the drying process, need to
be considered here.

It seems that it should be far more important to look at the areas where
brushstrokes were obviously painting ``rhythmically'': a series of fast strokes
arranged in a certain pattern, which can then be used to build up a personal
handwriting for a given artist\cite{johnson08image}.

Three different two-dimesnional wavelet transformation techniques were used in
this research:

\begin{itemize}
\item Penn State
\item Princeton
\item Maatrich
\end{itemize}

Penn State is an orthonormal series generated by a wavelet, implemented using
fast subband filtering, critically sampling the data.

Princeton forgoes critical sampling in favour of greater orientations using
complex wavelets.

Maatrich uses Gabor wavelets to provide a similar range of orientation
selectivity as Princeton, but in a way which is closer to a physiological
templates.

Because of the differences between these wavelet transforms, different
classifiers need to be applied to the resulting coefficients.

Both Penn State and Princeton transforms can use a hidden Markov model
approach. Because Markov models have states which are linked to the predecessor
or neighbour states with a probability distribution. Hidden Markov models are
Markov models where this probability distribution is unknown and can be
inferred from the data which constructs the model.

There are several classifiers which can be applied to Markov models, although
typically these relate to some form of expectation maximisation technique.

All experimental results were gathered from a data set containing 101 paintings
in the style of Van Gogh. 82 of which are believed to be true Van Gogh works,
13 which are currently in dispute and 6 which are known imitations. All of
these paintings are scaled using the same technique to be equivalent sizes.

The approach for using Penn State was to gather several unquestionably Van Gogh
paintings which are representative of his style (exemplar images) as a training
set. By not including negative examples in the training set the resulting
classifier is more stringent.

The paintings are then divided into patches of roughly $512 \times 512$ pixels
such that all the patches from the same painting are the same size.
Dissimilarity measures between these patches are calculated using both texture-
and stroke-based features. Paintings are then compared using the aggregated
distances between the patches of the images.

Texture features were extracted in a form which shows the abruptness of changes
in variations at different orientations. Textures were only captured at the
highest scale. These are then put into a feature vector which can be converted
into a Markov model.

For extraction of strokes an edge-detection algorithm is used to detect
contours of the image, which can then have several geometric features computed,
such length, average curvature and orientation.

Because textures are dependent on their neighbours it is sensible to build up a
probabilistic model, namely a Markov model, to show these dependencies. Because
the probabilities which affect the dependencies are unknown this must therefore
be a hidden Markov model, using two dimension (as images are two-dimensional).
A likelihood classifier can then be trained from the training set and applied
to other 2D hidden Markov models to be able to classify examples.

Spatial features are less important for brushstrokes as they capture more
global information rather than information which is likely to be affected by
neighbouring strokes. Clustering can be applied to this to partition the data
set then new examples can be classified using the distances to these clusters.

Princeton shows the scale at which details start to emerge; each patch is
analysed using the wavelet transform and then these coefficients are stored is
a hidden Markov tree in a similar process to \cite{jafarpour2009stylistic},
where the hidden states take the form of edge or non-edge. This helps show the
scale at which an individual artist's style emerges as the transitions between
edge and non-edge will occur at different scales.

Only 79 of all 101 paintings were analysed by Princeton, the other paintings
were too dark for meaningful analysis to be gained from them in this
experiment. 65 of these were van Goghs, 6 imitations and 5 of the questioned
paintings.

\gls{MDS} is used to find an arrangement of all the points in a three
dimensional space (as opposed to the very high dimensional space the hidden
Markov tree uses. For all van Gogh paintings the centre point is determined,
where all of these paintings are equally weighted.

The 3D visualisation shows a good separation of the van Goghs from the non-van
Goghs, but to be able to truly use this, a radius classifier is applied such
that a lower radius is more likely to classify as a van Gogh and visa versa. An
appropriate point $r$ is chosen as the cut distance at which the label changes,
such at van Gogh paintings are closer to the centre than $r$.

This classifier was tested using leave-one-out cross validation; a validation
technique based on $k$-fold cross validation where $k$ is the size of the data
set. $k$-fold cross validation partitions the data set into $k$ subsets. Each
subset has classification performed on its elements based on training from the
other subsets.

Leave-one-out cross validation is a useful cross-validation technique for small
data sets and can be used to gather a correlation of expected output against
classified result. This can also be used to gather the statistical significance
of the results.

Biologically inspired techniques have shown promising in other fields of
computer science, notably items like Artificial Neural Networks, Immune and
Endocrine Systems; Swarm Intelligence such as Ant Colony Optimisation; etc.

Using the human visual system as a metaphor paintings can be analysed using
three principals: 
\begin{inparaenum}[1)]
\item Importance of contours;
\item Analysis at multiple scales;
\item Similarities are reflected in local textures.
\end{inparaenum}

The Maatrich transform uses Gabor wavelet filters at multiple scales and
orientations. The provides an ``energy'' value for each pixel, orientation and
scale. These values can then be binned into a high-dimensional histogram. This
allows a \gls{SVM} to be trained and new example to be classified
with. Because there are only two states (van Gogh and non-van Gogh), a
\gls{SVM} is a good classifier. Again, leave-one-out cross-validation is used
for each painting, the patches are separately classified then the most
frequently returned label is used to classify the painting.

% PS results
Penn State begins to show some good results, although one of the imitations is
considered to be one of the five closest to the training set of 23 van Goghs.

% Pr results
The radius classifier applied to the Princeton tranformation successfully
classified 55 of 65 van Goghs and 9 out of 11 non-van Goghs correctly. At even
finer scale shows even better results.

% MA results
Maatrich transforms with a \gls{SVM} classifier successfully classified four of
the six non-van Goghs correctly, but wrongly classifying two of the van Goghs.
This starts to show that it can detect dissimilarities in the brushstroke
texture of a painting, but that some difference may be too subtle to pick up.

% Conclusions
The use of leave-one-out cross-validation means the experimental results are
very valid; but the lack of significance does remove some of this validity - it
is up for debate whether a null hypothesis could also be to play.

The images are high resolution greyscale, so although colour information has
been lost but the resolution means that most of the paintings should be uniform
and detailed.

The discussion and use of expert knowledge means that the results gained are
potentially more valid by not considering several sections of a painting.
However, the opposite argument could be made that part of van Goghs style was
to deform the paintings slight by leaning against them, etc. and therefore is a
better indicator of his style than some of the brushstrokes.

This does show positive results and the techniques are well thought out,
including discussions of other classifiers which could have been used.

As is a common problem in brushstroke extraction there is no ground truth on
which these techniques can be applied to, so it is difficult to know how well
these techniques will apply to other paintings by van Gogh or by different
artists.

In conclusion, this is a decent application of existing research with real
results being gained, although it doesn't deal with the extraction of
brushstrokes formally it is a good step in the right direction.

@@

\subsection{Stroke Decomposition}
% Paper 15

\subsection{Calligraphy Stroke Extraction}
% Paper 14
